% Place acknowledgments below.
Foremost, I am very grateful and fortunate for the support and guidance from my adviser, and dear friend, Dr. Michael Krane. His willingness to share his time, experience, and instruction to ensure my professional development has been invaluable. I would like to thank my previous mentors Dr. Timothy Brungart and Brian Kline for a) the memorable years of working together and b) their advice and encouragement continue my education. A special acknowledgement to Dr. Kofi Adu for introducing me to research, and shaping my thought process for problem solving. I would like to thank my committee members Dr. William Hancock and Dr. Adam Nickels for their oversight of my thesis and program. To our collaborators Dr. Justin Jaworski and Huansheng Chen from Lehigh University, I am very appreciative of the contributions, comments, and insight into the theoretical aspects of the experiment. Much of this would not have been possible without you. To my fellow friends and colleagues at the Garfield Thomas Water Tunnel Building, thank you for the continued support that you have given over the many years prior to \& during my master's program. I am grateful for the financial support received from the National Science Foundation (NSF CBET-180445) and the Applied Research Laboratory, making this possible. Finally, I would like to thank my friends, family, and loved ones for supporting me -- specifically, my dear friends Dr. Gage Walters \& Paul Trzcinski.

Any opinions, findings, and conclusions or recommendations expressed in this material are those of the author and do not necessarily reflect the views of the National Science Foundation.
